Essays in Computational Management Science

### 2018-08-05 Chapters/Chapter01.tex

The concept of transferring money using cryptography as an underlying technology was shortly presented in 1983 by \citet{chaum1983blind} and was deepened in a theoretical paper in 1985 \citep{chaum1985security}. However, it was only in 1988 that \citet{chaum1988untraceable} created the term \emph{electronic cash} and also proposed a basic and practical scheme which yielded untraceability yet allowed to trace double spendings.

@@

The network must scale to support hundreds of transactions per second, while its capacity is around only eight transactions per second. Thus, the more Bitcoin becomes popular, the more saturated the network is. Network saturation has many side effects and may affect the players' incentive to keep the network running. The transaction fees have to be increased to compete for faster confirmation. The pool of unconfirmed transactions grows indefinitely, which may cause some transactions to be discarded due to low memory space available, as the previously predictable confirmation time of transactions becomes unpredictable.

@@

Bitcoin seems to have the most decentralized network among the cryptocurrencies, even so, there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power (for details, see \citet{gencer2018decentralization}). Hence, they have an oversized influence when it comes to changes in the Bitcoin protocol's behavior. They may also cooperate in an attack, voiding transactions which seemed confirmed. The more decentralized, the more trustworthy Bitcoin is. This centralization problem is seen as an important challenge.

Generating new transactions in Bitcoin has a tiny computational cost because one only has to generate the transaction itself, digitally sign it, and propagate it in the Bitcoin network. On the one hand, it means that any device is capable of generating new transactions, but, on the other hand, it makes Bitcoin susceptible to spam attacks. One may generate hundreds of thousands of new valid transactions, overloading the unconfirmed transactions pool and saturating the network. This spam problem has happened several times and affects Bitcoin's trustworthy. \citet{bitcoinspam2017} reports a possible spam attack lasting at least contiguous 18 months.

@@

After the launch of Bitcoin, more than 1,000 other cryptocurrencies have been created \citep{coinmarketcap}. In general, they are Bitcoin-like, which means they use similar technologies, including the blockchain. Some cryptocurrencies differs a lot from Bitcoin, like the ones which use the Directed Acyclic Graph (DAG) model \citep{dagdiscussion2014, tangle2016, dagcoin2015, sompolinsky2013, lewenberg2015, vorick2015}. We are especially interested in one of them: Iota.

Iota uses a DAG model, called tangle, which has a different design than Bitcoin's blockchain. It has neither mining nor confirmation blocks and transaction fees. Each transaction has its own proof-of-work\footnote{The mechanism that assures the immutability is the proof-of-work, which makes it computationally infeasible to tamper with transactions. It will be explained later in details.} and is used to confirm other transactions, forming a directed acyclic graph of transactions. Thus, a transaction is said to be confirmed when there is enough proof-of-work from the transactions confirming it directly or indirectly. There is no other way to confirm transactions but generating new transactions.

In Iota, as transactions confirm transactions, the network benefits from a high volume of new transactions. Therefore, theoretically, it scales to any large number of transactions per second. The scaling problem of tangle is exactly the opposite of Bitcoin's: it must have at least a given number of transaction per seconds; otherwise, the transactions are not confirmed, and the cryptocurrency does not work. While Iota's network has not reached this minimum number of transactions per second, it uses a central coordinator which works as a trustworthy node \citep{iotacoordinator}.

Every transaction confirmed by the central coordinator is assumed to be valid and cannot be reverted. The remaining of the network can verify a confirmation through the central coordinator's digital signature. The coordinator will not be necessary anymore when the number of transactions per second reaches a minimum value, but Iota's developers cannot say precisely what is this minimum value. This just elucidates that the tangle does not seem to work properly under a low volume of transactions (and fluctuations in the number of transactions per second may severely affect Iota's trustworthiness).

The present work intends to propose and analyze a new architecture, named Hathor, which lies between Bitcoin and Iota and may be a viable solution to both scaling, centralization, and spam problems. We also present a mathematical analysis of Bitcoin's architecture.

@@

In \citeauthor{nakamoto2008bitcoin}'s (2009) seminal paper, there is no distinction between bitcoin and blockchain. They are just one thing which solves an important theoretical problem: how to create a distributed and decentralized digital form of hard money on the internet, in which all users can agree as to whom is entitled to which funds.

@@

The blockchain technology works through the creation of new blocks. Each new block confirms that all previous blocks are valid and have not been tampered with. The mechanism that assures the immutability is the proof-of-work, which makes it computationally infeasible to tamper with previous transaction records without having to recalculate all the previous proof-of-works faster than all of the remaining machines of the network. The network agrees that work should be done in the block at the longest chain in the blockchain.

@@

Bitcoin's blockchain uses the mathematical problem of finding a random number which, after being applied to the hash function SHA-256 twice, results in a number smaller than a given threshold $A$. As SHA-256 is a pseudo-random function, its output is uniformly distributed between 0 and $2^{256}-1$ \citep{gilbert2003security}. Thus, if the given number is $A = 2^{255}$, one has probability 50\% of finding a solution (just the most significant bit of the hash needs to be zero). But if the given number is $A = 2^{240}$, one has probability 0.0015\% of finding a solution (as the ~16 most significant bits of the hash must equal zero). Hence, finding a solution is a hard problem which difficulty depends on the given number $A$. The lesser the given threshold $A$, the higher the difficulty. On the other hand, checking whether a solution is correct is fast since one just has to apply the SHA-256 twice and compare.

@@

It may happen that two miners find two different blocks in a small interval of time. In this case, miners will propagate their blocks and part of the network will choose one of them as the next block, whereas the other part of the network will choose the other one. This phenomenon is called a fork. Thus, when the next block is found, one of those blocks will be confirmed, and the other will be ignored and referred to as an orphan block. As blocks confirm transactions, all the transactions in the orphan block which have not been confirmed by another block will return to the unconfirmed transaction pool. Hence, it is not safe to accept a transaction when it is confirmed by only one block. It is the idea behind the rule of thumb of waiting for at least six confirmations before accepting a transaction.

@@

A practical example of a network disagreement is the increase of the block size. No group with more than 50\% of the network's hash power agreed into increasing the maximum block size to increase the number of transactions confirmed by a block and thus increasing the network's capacity. Hence, the capacity remains the same, and the community has been discussing the issue in search of a consensus. For further details of the discussion, see \citet{bitcoinblocksize}.

@@

The digital signature scheme used by Bitcoin is based on a pair of private and public keys. The private key is used to sign the transaction, while the public key is used to check whether the signature is valid. Thus, the owner of some Bitcoin funds is, in fact, the owner of a pair of private and public keys. The private key must never be publicly published, as whoever has access to the private key is able to spend its funds. In other words, in order to protect their funds, the owners must protect the private key. If one loses their private key, unfortunately, access to their funds will be lost forever. The public key may be used to a proof-of-ownership, i.e., one may publish the public key with some message digitally signed by the private key, proving that he/she is the owner of the funds.

@@

Each transaction's output has a script that is executed by the miners to check whether one has or has not permission to spend that output. In other words, whether one has the ownership of that output. In order to execute these scripts, the miners also need some data. This data is given by the transaction which is spending the output.

The output's scripts usually checks whether the public key is valid and whether the digital signature was signed by the private key associated to that public key. Although there are only 3 commonly used scripts, one may create a custom script using the Bitcoin's script language \footnote{Your courageous author once tried to make a transaction with custom script to try to double spend a deposit to an exchange, only to learn through this intrepid adventure that Bitcoin allows only 3 script patterns and the others are treated as invalid.}.

The input contains the data which prove that the sender is the owner of the referred outputs, i.e., the input which must be accepted by the output scripts being spent. Usually, each input has the public key of the sender and a digital signature.

Those users accustomed to block explorers may have been misled by the transaction information that these websites provide.  For example, suppose a miner receives a transaction with ``one input from address A1''.  This ``one input'' actually consists of a pointer to a previous unspent output (i.e., there is no ``input'' address, as is displayed, but only a pointer). This pointer reference allows lookup to be executed in $O(1)$ time.

@@

Iota's underlying technology is Tangle, which has DAG-based architecture with a whole different approach to confirmations. It proposes that there is no need for a block to confirm transactions, as transactions can confirm themselves. Here, each transaction has its own proof-of-work, named weight, and they must confirm two other previous transactions. In this sense, instead of a chain of blocks, the transactions and their confirmations form a directed and acyclic graph (DAG), as in Fig. \ref{fig-tangle-example}.

Like Blockchain, Tangle is another technology to store immutable data and may be the underlying technology to different applications, such as cryptocurrencies, digital contracts (Ethereum-like), digital notaries, and so forth.

Transactions may be either confirmed or unconfirmed. The confirmed transactions have been already confirmed by at least one more transaction. It does not mean they are already irreversible and protected against a double spend attack --- it just means at least one transaction has done some work to confirm it. The unconfirmed transactions are called tips and they are eager to be confirmed. Usually a new transaction selects two tips to confirm, but this rule may not be followed.

Different from Bitcoin, transactions does not have scripts to check whether one may spend the tokens. Instead, it uses a fixed Winternitz hash-based digital signature \citep{dods2005hash}, i.e., whoever correctly signs the transaction may tranfer the tokens. It also supports multi-signature scheme (see \cite{iotamultisign}). Not allowing scripts is a disadvantage when compared to Bitcoin because future applications are limited and changing it would require a major modification in transaction format.

At the beginning, the digital signature algorithm relied on the Curl hash function, which is a ternary hash function designed by Iota's developers. This hash function was replaced by the Kerl hash function after \cite{heilman2017iota} has found a critical vulnerability which enabled practical signature forgery attacks. The Kerl hash function is a variation of SHA-3 also designed by Iota's developers \citep{iotakerl}. Like the Curl hash function, Kerl has not been deeply studied by cryptography researchers and may have critical vulnerabilities.

Iota's tokens are pre-mined, which means they have been issued in the genesis transaction and no more tokens will ever be issued. This means that there are no miners in Iota's network and only the users keep the network alive through their new transactions. This may lead to some incentive issues, because, on the one hand, users would need to keep transferring tokens to keep the network alive, while, on the other hand, users would need to believe that Iota is a good choice to keep using it. As users perception of quality is significantly affected by ``price and congestion'', the \emph{empty-restaurant syndrome} may keep new users away, making it even harder to reach the minimum required to work properly \citep{debo2010prices}. Even if the minimum has been reached, if the number of transactions per second plummet, the time to confirm the most recent transactions will increase significantly. This possibly poses a serious risk to the network stability and users' trustworthy.

As there are no miners, there are also no fees, which is a major incentive to newcomers. Tokens may be freely transferred without any losses, even for tiny amounts, enabling micropayments. For example, online workers paid per hour may receive their payments every hour, reducing the default risk. This allows untrusted parties to work together reducing the risk for both. The contractor always wants to pay in the end, while the workers always wants to get paid in the beginning.

The Internet-Of-Things (IoT) technologies, which is a network of smart devices that are connected to the internet and exchange data, also benefits from not having fees. Imagine your refrigerator doing your groceries and automatically paying using your Iota tokens, or your electric car automatically paying for recharge. In both situations, transactions with no fees are interesting because these devices can freely make many microtransactions without any loss. For further information, see \cite{fleisch2010internet}.

Theoretically, Iota's network benefits from high volume of transactions. The more transactions are coming, the faster previous transactions are confirmed. This is a major feature of Tangle which contrasts with Bitcoin's difficulty to scale. On the other hand, low volume of transactions is a primary problem for Iota, since it would take too long to confirm previous transactions. As it is a known problem, Iota has created a central coordinator which works as a trustworthy node, clearing transactions. In other words, every transaction directly or indirectly confirmed by the central coordinator is assumed to be cleared. Iota's developers claims that the coordinator will be turned off when the network outgrows a minimum (unknown) size.

An important factor of Iota is how the new transactions choose which transactions they will confirm. There are several possible approaches, such as randomly selecting two of the unconfirmed transactions (tips). In Fig \ref{fig-tangle-example}, the reader may have noticed that transaction 8 will confirm transaction 4, which has already been confirmed by transaction 7. It may be on purpose, or maybe transaction 4 was unconfirmed when it was chosen, but it got confirmed during the calculation of the proof-of-work or the network propagation of the transaction. The selection algorithm seems to be important to protect the network against double spend attacks. For further information, see \cite{tangle2016}.

>>>>

The concept of transferring money using cryptography as an underlying technology was briefly presented in 1983 by \citet{chaum1983blind} and was deepened in a theoretical paper in 1985 \citep{chaum1985security}. However, it was only in 1988 that \citet{chaum1988untraceable} created the term \emph{electronic cash} and also proposed a basic and practical scheme which yielded untraceability yet allowed to trace double spendings.

@@

% Eric: why does it must scale?
The network must scale to support hundreds of transactions per second, while its capacity is around only eight transactions per second. Thus, the more Bitcoin becomes popular, the more saturated the network is. Network saturation has many side effects and may affect the players' incentive to keep the network running. The transaction fees have to be increased to compete for faster confirmation. The pool of unconfirmed transactions grows indefinitely, which may cause some transactions to be discarded due to low memory space available, as the previously predictable confirmation time of transactions becomes unpredictable.

@@

Bitcoin seems to have the most decentralized network among the cryptocurrencies, even so, there are a few miners and mining pools which together control over 50\% of the network’s computing (hash)power (for details, see \citet{gencer2018decentralization}). Hence, they have an oversized influence when it comes to changes in the Bitcoin protocol's behavior. They may also cooperate in an attack, voiding transactions which seemed confirmed. The more decentralized, the more trustworthy Bitcoin is. This centralization problem is seen as an important challenge.

Generating new transactions in Bitcoin has a tiny computational cost because one only has to generate the transaction itself, digitally sign it, and propagate it in the Bitcoin network. On the one hand, it means that any device is capable of generating new transactions, but, on the other hand, it makes Bitcoin susceptible to spam attacks. One may generate hundreds of thousands of new valid transactions, overloading the unconfirmed transactions pool and saturating the network. This spam problem has happened several times and affects Bitcoin's trustworthiness. \citet{bitcoinspam2017} reports a possible spam attack lasting at least a contiguous 18 months.

@@

Since the launch of Bitcoin, more than 1,000 other cryptocurrencies have been created \citep{coinmarketcap}. In general, they are Bitcoin-like, which means they use similar technologies, including the blockchain. Some cryptocurrencies differ a lot from Bitcoin, like the ones which use the Directed Acyclic Graph (DAG) model \citep{dagdiscussion2014, tangle2016, dagcoin2015, sompolinsky2013, lewenberg2015, vorick2015}. We are especially interested in one of them: Iota.

Iota uses a DAG model, called Tangle, which has a different design than Bitcoin's blockchain. It has neither mining nor confirmation blocks and transaction fees. Each transaction has its own proof-of-work\footnote{The mechanism that assures the immutability is the proof-of-work, which makes it computationally infeasible to tamper with transactions. It will be explained later in details.} and is used to confirm other transactions, forming a directed acyclic graph of transactions. Thus, a transaction is said to be confirmed when there is enough proof-of-work from the transactions confirming it either directly or indirectly. There is no other way to confirm transactions other than generating new transactions.

In Iota, as transactions confirm transactions, the network benefits from a high volume of new transactions. Therefore, theoretically, it scales to any large number of transactions per second. The scaling problem of Tangle is exactly the opposite of Bitcoin's: it must have at least a given number of transactions per second; otherwise, the transactions are not confirmed, and the cryptocurrency does not work. When Iota's network has not reached this minimum number of transactions per second, it uses a central coordinator that works as a trustworthy node \citep{iotacoordinator}.

Every transaction confirmed by the central coordinator is assumed to be valid and cannot be reverted. The remainder of the network can verify a confirmation through the central coordinator's digital signature. The coordinator will not be necessary anymore when the number of transactions per second reaches a minimum value, but Iota's developers cannot say precisely what is this minimum value. This just elucidates that the Tangle does not seem to work properly under a low volume of transactions (and fluctuations in the number of transactions per second may severely affect Iota's trustworthiness).

The present work intends to propose and analyze a new architecture, named Hathor, which is situated on a spectrum between Bitcoin and Iota and may be a viable solution to scaling, centralization, and spam problems. We also present a mathematical analysis of the Bitcoin's architecture.

% TODO Eric: Why Bitcoin? Not Hahor?

@@

In \citeauthor{nakamoto2008bitcoin}'s (2009) seminal paper, there is no distinction between bitcoin and blockchain. They are just considered parts of a single entity which solves an important theoretical problem: how to create a distributed and decentralized digital form of hard money on the internet, in which all users can agree as to whom is entitled to which funds.

@@

The blockchain technology works through the creation of new blocks. Each new block confirms that all previous blocks are valid and have not been tampered with. The mechanism that assures the immutability is the proof-of-work, which makes it computationally infeasible to tamper with previous transaction records without having to recalculate all the previous proofs-of-work faster than all of the remaining machines of the network. The network agrees that work should be done in the block at the longest chain in the blockchain.

@@

Bitcoin's blockchain uses as proof-of-work the mathematical problem of finding a random number which, after being applied to the hash function SHA-256 twice, results in a number smaller than a given threshold $A$. As SHA-256 is a pseudo-random function, its output is uniformly distributed between 0 and $2^{256}-1$ \citep{gilbert2003security}. Thus, if the given threshold is $A = 2^{255}$, one has probability 50\% of finding a solution (just the most significant bit of the hash needs to be zero). But if the given number is $A = 2^{240}$, one has probability 0.0015\% of finding a solution (as the ~16 most significant bits of the hash must equal zero). Hence, finding a solution is a hard problem which difficulty depends on the given number $A$. The lesser the given threshold $A$, the higher the difficulty. On the other hand, checking whether a solution is correct is fast since one just has to apply the SHA-256 twice and compare.

@@

It may happen that two miners find two different blocks in a small interval of time. In this case, miners will propagate their blocks and part of the network will choose one of them as the next block, whereas the other part of the network will choose the other one. This phenomenon is called a fork. Thus, when the next block is found, one of those blocks will be confirmed, and the other will be ignored and referred to as an orphan block. As blocks confirm transactions, all the transactions in the orphan block which have not been confirmed by another block will return to the unconfirmed transaction pool. Hence, it is not safe to accept a transaction when it is confirmed by only one block. This idea is behind the rule of thumb of waiting for at least six confirmations before accepting a transaction.

@@

A practical example of a network disagreement is the increase of the block size. In previous attempts to do so, no group with more than 50\% of the network's hash power agreed to increasing the maximum block size to increase the number of transactions confirmed by a block with the goal of thus increasing the network's capacity. Hence, the capacity remains the same, and the community has been discussing the issue in search of a consensus. For further details of the discussion, see \citet{bitcoinblocksize}.

@@

The digital signature scheme used by Bitcoin is based on a pair of private and public keys. The private key is used to sign the transaction, while the public key is used to check whether the signature is valid. Thus, the owner of some Bitcoin funds is, in fact, the owner of a pair of private and public keys. The private key must never be publicly published, as whoever has access to the private key is able to spend its funds. In other words, in order to protect their funds, the owners must protect the private key. If one loses their private key, unfortunately, access to their funds will be lost forever. The public key may be used on a proof-of-ownership, i.e., one may publish the public key with some message digitally signed by the private key, proving that he/she is the owner of the funds.

@@

Each transaction's output has a script that is executed by the miners to check whether one has or has not permission to spend that output. In other words, the script verifies whether one has the ownership of that output. In order to execute these scripts, the miners also need some data. This data is given by the transaction which is spending the output.

The output's scripts usually check whether the public key is valid and whether the digital signature was signed by the private key associated with that public key. Although there are only 3 commonly used scripts, one may create a custom script using the Bitcoin's script language \footnote{Your courageous author once tried to make a transaction with custom script to try to double spend a deposit to an exchange, only to learn through this intrepid adventure that in practice Bitcoin allows only 3 script patterns and the others are treated as invalid.}.

The input contains the data which proves that the sender is the owner of the referred outputs, i.e., the input which must be accepted by the output scripts being spent. Usually, each input has the public key of the sender and a digital signature.

Those users accustomed to online block explorer websites may have been misled by the transaction information that these websites provide.  For example, suppose a miner receives a transaction with ``one input from address A1''.  This ``one input'' actually consists of a pointer to a previous unspent output (i.e., there is no ``input'' address, as is displayed, but only a pointer). This pointer reference allows lookup to be executed in $O(1)$ time.

@@

Iota's underlying technology is Tangle, which has a DAG-based architecture with a wholly different approach to confirmations. It proposes that there is no need for a block to confirm transactions, as transactions can confirm themselves. Here, each transaction has its own proof-of-work, named weight, and must confirm two other previous transactions. In this sense, instead of a chain of blocks, the transactions and their confirmations form a directed and acyclic graph (DAG), as in Fig. \ref{fig-tangle-example}.

Like Blockchain, Tangle is another technology to store immutable data and may be the underlying technology for different applications, such as cryptocurrencies, digital contracts (like Ethereum), digital notaries, and so forth.

Transactions always exist in one of two states: confirmed or unconfirmed. The confirmed transactions have been already confirmed by at least one other transaction. Being confirmed does not mean that a transaction is already irreversible and protected against a double spending attack --- it just means at least one transaction has done some work to confirm it. The unconfirmed transactions are called tips and they are eager to be confirmed. Usually a new transaction selects two tips to confirm, but this rule is not enforced.

In contrast to Bitcoin, transactions do not have scripts to check whether one may spend the tokens. Instead, Iota uses a fixed Winternitz hash-based digital signature \citep{dods2005hash}, i.e., whoever correctly signs the transaction may transfer the tokens. It also supports a multi-signature scheme (see \cite{iotamultisign}). Not allowing scripts is a disadvantage when compared to Bitcoin because future applications are limited and changing Iota would require a major modification in transaction format.

At the beginning of Iota's development, the digital signature algorithm relied on the Curl hash function, which is a ternary hash function designed by Iota's developers. This hash function was replaced by the Kerl hash function after \citet{heilman2017iota} has found a critical vulnerability that enabled practical signature forgery attacks. The Kerl hash function, also designed by Iota's developers \citep{iotakerl}, is a variation of SHA-3. Like the Curl hash function, Kerl has not been deeply studied by cryptography researchers and may have critical vulnerabilities that remain unknown to the public.

Iota's tokens are pre-mined, which means they have all been issued in the genesis transaction and no more tokens will ever be issued. This means that there are no miners in Iota's network and only the users keep the network alive through their new transactions. This may lead to some incentive issues, because, on the one hand, users would need to keep transferring tokens to keep the network alive, while, on the other hand, users would need to believe that Iota is a good choice to keep using it. As users' perception of quality is significantly affected by ``price and congestion'', the \emph{empty-restaurant syndrome} may keep new users away, making it even harder to reach the minimum required to work properly \citep{debo2010prices}. Even if the minimum has been reached, if the number of transactions per second plummets, the time to confirm the most recent transactions will increase significantly. This possibly poses a serious risk to the network stability and users' trust.

As there are no miners, there are also no fees, which is a major incentive to newcomers. Tokens may be freely transferred without any losses, even for tiny amounts, enabling micropayments. For example, online workers paid per hour may receive their payments every hour, reducing their default risk. This allows untrusted parties to work together, with reduced risk for both. The contractor always wants to pay at the end of a job, while the workers always want to get paid in the beginning.

The Internet-Of-Things (IoT) technologies, which comprise a network of smart devices that are connected to the internet and exchange data, also benefits from not having fees. Imagine your refrigerator doing your groceries and automatically paying using your Iota tokens, or your electric car automatically paying for recharge. In both situations, transactions with no fees are interesting because these devices can freely make many microtransactions without any loss. For further information, see \cite{fleisch2010internet}.

Theoretically, Iota's network benefits from a high volume of transactions. The more transactions are coming, the faster previous transactions are confirmed. This is a major benefit of Tangle, which contrasts with Bitcoin's difficulty to scale. On the other hand, low transaction volume is a primary problem for Iota, since it would take too long to confirm previous transactions. As when few transactions arrive is a known problem, Iota has created a central coordinator which works as a trustworthy node, clearing transactions. In other words, every transaction directly or indirectly confirmed by the central coordinator is assumed to be cleared. Iota's developers claim that the coordinator will be turned off when the network outgrows a minimum \emph{unknown} size.

An important factor of Iota is how the new transactions choose which transactions they will confirm. There are several possible approaches, such as randomly selecting two of the unconfirmed transactions (tips). In Fig \ref{fig-tangle-example}, the reader may have noticed that transaction 8 will confirm transaction 4, which has already been confirmed by transaction 7. It may have been on purpose, or perhaps transaction 4 was unconfirmed when it was chosen, but it got confirmed during the calculation of the proof-of-work or the network propagation of the transaction. The selection algorithm seems to be critical in protecting the network against double spend attacks. For further information, see \cite{tangle2016}.
